Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-04 how.tex
Updates from ShareLaTeX

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
This paper presents
HOW, a novel  {\em instance-based planner} that runs in near-linear time.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item The software defects found in  twelve object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the six software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
Many papers use this above data to build {\em predictors} for defects and runtimes.
However, aside from work by the authors of this paper,
none of those papers then go on to use this data to build {\em planners}.

In summary, this paper makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner.
\item A proof that HOW runs in near-linear time.
\item An experimental validation that the HOW planner is useful for reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\ee
The rest of this paper is structured as follows.
HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. As discussed in \tion{mm},
this is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}.

@@

 \subsectuion{Why ``HOW''?}

HOW was inspired by some recent work by Krall and Menzies and Davis on 
the GALE optimizer~\cite{krall14}. A well-known effect in machine learning (and, in computer vision
and compression~\cite{levina04}) is that the $n$ dimensions of a data set can be represented by a  much
lower-dimensional approximation. GALE finds those $m << n$ dimensions then mutates examples
towards the better end of each dimension $m_i$.

A limitation with GALE is that whenever it adjusts an instance, some model has to be called
to re-evaluate the new objective scores for the new examples. This can be a limitation
since, in some domains, no such model exists or is readily accessible.
Building and certifying a model   takes a long time. Such slow-to-build models exist in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).  This is not an isolated example. One of us (Menzies) teaches a graduate class in automated model-based software engineering. 
Students in that class are asked to find (or build) a non-trivial model of
some software process (where ``non-trivial'' means something that processes
a  model at least an order of magnitude more complex than Brooks' Law). 
While the  number of such models has been growing in recent years\footnote{See the
catalogue at goo.gl/nv2AVK.}, 
the total number remains very small-- which is an observation consistent with
SE models typically require  much effort to build and certify.

Finally, even when there is an existing model, instance-based reasoning can be useful.
Sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

Depending on the audience, we can
might call HOW a spectral learner, 
a response surface method, or  a local search operator.
{\em Spectral learners}~\cite{kamvar03} reason across $m$ dimensions synthesized
using (say) PCA~\cite{pearson1901}. For example,
algorithms like PDDP~\cite{boley98} use PCA to
recursively divide data into smaller regions where each level of recursion needs an   $O(N^2)$ analysis
to infer the principle components. HOW performs an analogous procedure using a   $O(2N)$ analysis
based on a heuristic proposed by
Faloutsos and Lin~\cite{Faloutsos1995} (see above, the ``FASTMAP'' algorithm).

{\em Response surface methods}  build a fast approximation to the function being optimized,
then uses that  approximation to take intelligent guesses about useful mutations. HOW
is a non-parametric response surface method that finds its response surface via recursive
clustering. For an example of a parametric response surface method, that uses Gaussian process
models, see Zuluaga et al.~\cite{zuluaga13}.

A {\em local search operator} that a some probability $p$, nudges  instance variables
along a dimension that improves its objective scores. HOW's recursive clustering methods
combine $n$ dimensions into $m \ll n$ so its ``nudges'' are simultaneous
adjustments along multiple dimensions. For an example of other kinds of local search algorithms, that only adjust
one variable at a time, see the literature on WALKSAT~\cite{Selman1992} and MAXWALKSAT~\cite{kautz97}.
See also the literature on local search methods in multi-objective optimization.

@@

HOW divide training into regions where the training
data has similar properties. Each region is then examined and divided into ``west'' and ``east''
where the region's ``west'' end has examples with worst performance stores than the east
end (mnemoc: west starts with ``w'' and is short for ``worst' and east starts with ``e;;]which is shirt
for excellent).so the ``e'' in

\be
\item Clustering the training data 
\item 
(c) finding which end of the cluster has better performance scores.
(d) find

@@

\section{Related Work}
Policy generation is different to standard data mining in that the 
latter is most concerned with what {\em is} rather than {\em what to do}.
For example, a classification or prediction system such as C4.5~\cite{quinlan92} or CART~\cite{breiman84} can predict for a discrete or continuous class variable
(respectively). As such, they can be used as a sub-routine within a
policy generator. For example, later in this paper we will use the CART
prediction system to assess the policies generated by HERE.

Policy generation with HERE is different to the recommendor systems.
As described by Robillard and Walker~\cite{rob14}, recommendor systems
in software engineering
are less {\em directive} and more    {\em collaborative} than policy
generators:
\bi
\item
Directive systems such as HERE tell developers exactly what to do
(e.g. ``use these settings in a Makefile'') 
\item
Collaborative systems are more suggestive guides that
help (e.g.) programmers focus on very small sets
with  very large libraries of code of documentation.
\ei
Collaborative systems may not presume to know what to do
in that smaller set (while the directive systems would issue an edict
such as  ``change this to that'').

We also distinguish HERE's policy generation fro the multi-objective optimization methods explored by Harman and Krall and Menzies and 
and many others.

Previous with have explored model-based policy generation with STAR,
NOVA, and instance-based policy generation with  W2.

refactoring XXX

Finally, we note that HERE is a temporal transfer learner.

@@

% \section{Discussion}
% \section{Threats to validity}
% \section{Conclusion}
% \section*{Acknowledgements}
\newpage
\bibliography{how}{}
\bibliographystyle{IEEEtran}
\end{document}

>>>>

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
This paper presents
HOW, a novel  {\em instance-based planner} that runs in near-linear time.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item The software defects found in  twelve object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the six software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
In summary, this paper makes six contributions that are significantly different to
prior work:
\be
\item Many software analytics papers use data to build {\em predictors} for (e.g.) defects and runtimes.
However, aside from work by the authors of this paper,
none of those papers then go further to use that data to build {\em planners}.
\item This paper offers a general framework for augmenting a prediction system such that
it becomes a planning system;
\item This paper offers a  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item This paper offers a general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner of this paper has not been presented before.
\item The experiments of this paper show that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\ee
The rest of this paper is structured as follows.
HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. As discussed in \tion{mm},
this is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}.

@@

 \subsection{Motivation: Why ``HOW''?}
HOW was inspired by some recent work by Krall and Menzies and Davis on 
the GALE optimizer~\cite{krall14}. A well-known effect in machine learning (and, in computer vision
and compression~\cite{levina04}) is that the $n$ dimensions of a data set can be represented by a  much
lower-dimensional approximation. GALE finds those $m << n$ dimensions then mutates examples
towards the better end of each dimension $m_i$. Since it explores $m \ll n$ dimensions,
GALE terminates much faster than standard optimiser (e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}).
Better yet, GALE's rapdily generated  solutions are competitive (or better) than those found by 
slower optimizers.

A limitation with GALE is that it makes extensive use of
an underlying domain model to asses generations of newly mutated instances.  This is an issue since
in some domains, no such model exists or is readily accessible. For example, this whole paper
began when, after briefing a business client on GALE, they asked:
\begin{quote}
{\em ``But what if I do not have a model of my business?''}
\end{quote}
This is a legitimate concern.
Building and certifying a model  can  takes  long time, especially in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).  This is not an isolated example. One of us (Menzies) teaches a graduate class in automated model-based software engineering. 
Students in that class are asked to find (or build) a non-trivial model of
some software process (where ``non-trivial'' means something that processes
a  model at least an order of magnitude more complex than Brooks' Law). 
While the  number of such models has been growing in recent years\footnote{See the
catalogue at goo.gl/nv2AVK.}, 
the total number remains very small-- which is an observation consistent with
SE models typically require  much effort to build and certify.

Also, even when there is an existing model, instance-based reasoning can be useful.
Models need maintenance lest they become out-dated. Elsewhere, we have described our
extensions to the USC models to enable reasoning about agile software developments. 
It took many months to implement and certify those extensions~\cite{me09i,me09j}.
The problem of model maintenance is another
motivation to look for alternate methods that can be automateically updated whenever new data arrives.

Finally, sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

Note that this lack of access to a workable model is not just fatal for GALE, but also for every
other model-based optimization algorithm such as simulated
annealling or evolutuionary methods such as   
NSGA-II, NSGA-III, SPEA2, IBEA, PSO, DE, MOEA/D, etc.~\cite{deb00a,zit02,zit04,%
deb14,Cui2005a,storn97,zhang07:TEC}. Accordingly, this paper seeks
an optimization methods that avoids the problem of models that are missing, slow
to build, hard to maintain, or difficult to access.

\subsection{Methods: How to Build ``HOW''?}
HOW is a variant of GALE that does not need a underlying domain model.

When designing HOW, our premises were:
\bi
\item
This is no domain model; (b)~we have some
labelled data and (b)~we seek minimal
disrupttion
HOW divide training into regions where the training
data has similar properties. Each region is then examined and divided into ``west'' and ``east''
where the region's ``west'' end has examples with worst performance stores than the east
end (mnemoc: west starts with ``w'' and is short for ``worst' and east starts with ``e;;]which is shirt
for excellent).so the ``e'' in

@@

\be
\item Clustering the training data 
\item 
(c) finding which end of the cluster has better performance scores.
(d) find
\ee

@@

@@

% \section{Discussion}
\section{Threats to validity}
\input{related}

\section{Conclusion}
% \section*{Acknowledgements}
\newpage
\bibliography{how}{}
\bibliographystyle{IEEEtran}
\end{document}
